---
title: "Depth Profiles"
author: "San Francisco Bay Study"
date: ""
output:
  powerpoint_presentation: default
  slidy_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(patchwork)
library(deltadata)
```

```{r, include=FALSE}
# Get color palette
library(Polychrome)

# build-in color palette
set.seed(76543) # this allows the result to be the same every time we run the code
P55 = createPalette(55,  c("#ff0000", "#00ff00", "#0000ff")) # these are our base colors
swatch(P55) # prints out a swatch of all the colors in the palette
names(P55) <- NULL # removes the names of the colors
```


```{r, include=FALSE}
# Define variables
# ID what survey we are looking at
year <- 2025
month <- "February"
survey = 2
currentDate <- Sys.Date()

# file path for lab entry file on u-drive
folder <- paste("U:\\LTM\\BayStudy_Lab\\", year, "_Data\\", sep = "")
file <- paste(year, "BS Field Entry.mdb", sep = " ")
fieldentry_path <- paste(folder, file, sep = "") # path for field entry file
```


```{r, include=FALSE}
# Get mate data for each station (sometimes we want to plot profiles by Captain); currently the profiles are not plotted by mate

# read in mate info
matetable<- bridgeAccess(paste(folder, "BSTrawlHistory.mdb", sep = ""),
             c("Main"))

# read in station/tow data
stationtable <- bridgeAccess(fieldentry_path,
                        c("Station", "StationConstants"))

# combine mate and station data 
station <- stationtable$Station %>% 
  select(Year, Survey, Station, Date) %>% 
  mutate(Date = as.POSIXct(Date))
bay <- stationtable$StationConstants %>% 
  select(Station, Bay)
mate_data <- matetable$Main %>% 
  select(Year, Survey, SamplingDate, Mate) %>% 
  rename(Date = SamplingDate) #%>% 
  #mutate(Date = as.character(Date))

# this is the mate data that we will join with the profile data next
mates <- station %>% 
  left_join(mate_data, by = c("Year", "Survey", "Date")) %>% 
  filter(Year %in% c(year)) %>% 
  rename(Month = Survey)
```


```{r, include=FALSE}  
# read in data from previous r markdown
folder_depthloggers <- paste("U:\\LTM\\Bay Study\\Depth Loggers\\Data\\", year, "\\", sep ="")
file_depthprofile <- paste(year, month, "WQprofiles_forplotting.csv", sep = "")
depthprofile_path <- paste(folder_depthloggers, file_depthprofile, sep = "")

# clean data 
Profiles <- read_csv(depthprofile_path) %>% # file we made in previous script
  mutate(MonthName = month.abb[Month]) %>% # making a column for an abbreviated month name (e.g., January -> Jan)
  filter(Month == survey) %>% # filter data for what survey we are working on, though data likely only include this survey anyway
  left_join(mates, by =c("Month", "Station", "Year")) %>% # join mate data
  mutate(BoatOpCode = ifelse(Mate %in% "Hull", "A", "B")) %>% # make a new column that IDs made by a letter rather than their name
  #filter(!Offset_Sec < DeployTime) %>% # remove depth readings that are before the actual set time
  #filter(!Offset_Min > 20) %>% # options to remove depth readings after a certain amount of minutes; this is usually when we are bringing the net in
  mutate(StationReplicate = as.numeric(paste(Station, replicate, sep = "."))) %>% # make a column with the station and replicate number if you want to plot a station with retows
  left_join(bay, by = c("Station")) %>% # join bay data - this allows you to split up the profiles by bay, so the plots are not super cluttered (mainly helps with cable out = 200 stations because we have a lot of those)
  #left_join(channeltows, by = c("Station", "replicate")) %>% # this is old, probably can delete
  filter(replicate == 1) %>% # use this filter if you only want good tows (e.g., no retows)
  filter(Depth < 50) %>%  # a quick fix for weird depth readings that are way too deep; these must be errors
  filter(Depth > 0) #another quick fix when the logger has negative reading (i.e., logger is above the water)
```

```{r, include=FALSE}
## The following chunks each produce a slide in a power point. The profiles are separated by the cable out distance.

## To create the powerpoint, click the 'Knit' button at the top of the window. The powerpoint will appear as a file in your working directory. Save this powerpoint here for 2023: U:\LTM\Bay Study\Depth Loggers\Data\2023\Plots. In the powerpoint, you can add comments on any weird profiles. For example, a tow could look weird, but according to the datasheet there was a lot of wind etc. Present the power point at the monhtly meetings as needed. 
```


## Cable Out = Bridles + 50

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  #filter(Depth < 7) %>% # use if there are any bad readings that go too deep
  filter(CableOut == 50) %>% 
  #filter(Station == 320) %>% # use if you want to look at one or a smaller amount of stations
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) + # assign the color level to Stations
  scale_color_manual(values = P55) + # assign colors
  scale_y_reverse() + # flip y axis so it looks like a depth profile (i.e., surface is on top)
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) 
```

## Cable Out = Bridles + 100

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  filter(CableOut == 100) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) 
```

## Cable Out = Bridles + 125

```{r, fig.width=20, fig.height=10, include = F}
Profiles %>%
  filter(CableOut == 125) %>%
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28))
```

## Cable Out = Bridles + 150

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  filter(CableOut == 150) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) 
```

## Cable Out = Bridles + 175

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  filter(CableOut == 175) %>% 
  filter(Offset_Min < 16) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) 
```

## Cable Out = Bridles + 200 - Delta 

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  filter(Bay %in% c(5:7)) %>% 
  filter(replicate == 1) %>% 
  filter(CableOut == 200) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) +
  guides(color = guide_legend(override.aes = list(size = 5))) 
  #+    facet_wrap(.~as.factor(Bay))
```

## Cable Out = Bridles + 200 - Suisun Bay

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  filter(Offset_Min < 18) %>% 
  filter(Bay %in% c(4)) %>% 
  filter(replicate == 1) %>% 
  filter(CableOut == 200) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) +
  guides(color = guide_legend(override.aes = list(size = 5))) +  
  facet_wrap(.~as.factor(Bay))
```

## Cable Out = Bridles + 200 - San Pablo Bay

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  #filter(Station == 345) %>% 
  #filter(Offset_Min < 18) %>% 
  filter(Bay %in% c(3)) %>% 
  filter(replicate == 1) %>% 
  filter(CableOut == 200) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) +
  guides(color = guide_legend(override.aes = list(size = 5))) 
  #+facet_wrap(.~as.factor(Bay))
```

## Cable Out = Bridles + 200 - Central Bay

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  #filter(Station == 345) %>% 
  #filter(Offset_Min < 18) %>% 
  filter(Bay %in% c(2)) %>% 
  filter(replicate == 1) %>% 
  filter(CableOut == 200) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) +
  guides(color = guide_legend(override.aes = list(size = 5))) 
  #+ facet_wrap(.~as.factor(Bay))
```

## Cable Out = Bridles + 200 - South Bay

```{r, fig.width=20, fig.height=10}
Profiles %>% 
  #filter(Station == 345) %>% 
  #filter(Offset_Min < 18) %>% 
  filter(Bay %in% c(1)) %>% 
  filter(replicate == 1) %>% 
  filter(CableOut == 200) %>% 
  ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
  geom_line(color = "black") +
  geom_point(aes(color = as.factor(Station)), size = 4) +
  scale_color_manual(values = P55) +
  scale_y_reverse() +
  labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 28)) +
  guides(color = guide_legend(override.aes = list(size = 5))) 
  #+  facet_wrap(.~as.factor(Bay))
```

```{r, fig.width=12, fig.height=6, include=FALSE}
## Cable Out = Bridles + 225
# Only use if we have a tow with 225 out. The file with not knit to a PDF if this is empty. Move the line above out of the chunk if we have tows at 225 out. Also, remove the phrase "include = FALSE" in the chunk title if we want this in the ppt.  
# Profiles %>%
#   filter(CableOut == 225) %>%
#   ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
#   geom_line(color = "black") +
#   geom_point(aes(color = as.factor(Station)), size = 3) +
#   scale_color_manual(values = palette) +
#   scale_y_reverse() +
#   labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
#   theme_bw() +
#   theme(legend.position = "bottom") +
#   #+ guides(color = guide_legend(override.aes = list(size = 5))) +
#   facet_wrap(~as.factor(MonthName))
```



```{r, fig.width=20, fig.height=10, include=FALSE}
## MWT towed in opposite direction
# Profiles %>% 
#   #filter(Bay %in% c(1:2)) %>% 
#   filter(replicate == 1) %>% 
#   filter(CableOut == 200) %>% 
#   #filter(Net == 1 & Direction == 2) %>% # dataset does not have direction right now
#   #filter(Station == 345) %>% # ID stations where we towed in the opposite direction
#   ggplot(aes(x=Offset_Min, y=Depth, group = Station)) +
#   geom_line(color = "black") +
#   geom_point(aes(color = as.factor(Station)), size = 4) +
#   scale_color_manual(values = P55) +
#   scale_y_reverse() +
#   labs(x="Tow Duration (Minutes)", y = "Depth (Meters)", color = "Station") +
#   theme_bw() +
#   theme(legend.position = "bottom") +
#   theme(text = element_text(size = 19)) +
#   guides(color = guide_legend(override.aes = list(size = 5))) +  
#   facet_wrap(.~as.factor(Bay))
```

Get data and clean to plot speed through the water
```{r, include=FALSE}
# read in file made in previous r markdown
STWdata <- read_csv(paste("U:\\LTM\\Bay Study\\Depth Loggers\\Data\\", year, "\\", year, "STW.csv", sep ="")) %>% 
  filter(Survey == survey) 

# calculate monthly average
monthavg <- mean(STWdata$CalcSTW_mpersec, na.rm = T)
```

Plot speed through the water

*red lines are our target range
*dashed blue is the survey average
```{r, fig.width=20, fig.height=10, message=FALSE, warning=FALSE}
# plot STW with monthly average. Red lines represent the target range
STWdata %>% 
  ggplot() +
  geom_point(aes(x=as.factor(Station), y=CalcSTW_mpersec), shape = 21, size = 6, fill = "grey") +
  geom_hline(yintercept = 0.67, color = "red", size = 1) +
  geom_hline(yintercept = 0.95, color = "red", size = 1) +
  geom_hline(yintercept = monthavg, color = "cyan", linetype = 2, size = 1) +
  xlab("Station") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 24),
        axis.text.y = element_text(size = 28)) +
    theme(text = element_text(size = 28))

```

